---
title: "Grants data frame"
output: github_document
html_document:
  df_print: paged
  keep_md: true
  code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( message=F, warning=F, fig.width = 10, fig.height = 10 )
library(dplyr)
library(pander)
setwd("C:/Users/icps/Dropbox/3. Maxwell/4. Capston/Data-Capstone/ignacio")
gra16 <- readRDS( gzcon(url("https://github.com/DataCapstone/Data-Capstone/blob/master/Raw-Data/NYgra16.rds?raw=true")))


```


```{r, eval=FALSE}
#generating a csv with the variables
x <- gra16[1,]
x <- as.data.frame(lapply(x, class))
x <- as.data.frame(t(x))
x <- data.frame(var = row.names(x), class = as.character(x$V1), stringsAsFactors = F)
grav <- x
write.csv(grav, "gra.variables.csv", row.names = F)

```

```{r, eval= FALSE}
#exploring some variables

#types of assistance
#what type of assistances are they getting?
unique(gra16$asst_cat_type)
unique(gra16$assistance_type)

#cfda_program_title
head(gra16$cfda_program_title)
head(gra16$cfda_program_num)
(table(gra16$cfda_program_title == "")) #has 110 blank ones.
table(is.na(gra16$cfda_program_title)) #no NAs here!
class(gra16$cfda_program_title)

#fda_program_num
table(is.na(gra16$cfda_program_num)) #no NAs here!
table(gra16$cfda_program_num == "") #no blanks!
class(gra16$cfda_program_num)

#project_description
unique(gra16$project_description)

######PRINCIPAL PLACE###############
#location: wanna focus on 4 cities, no idea if they are comparable.
sort(unique(gra16$principal_place_cc))
# "Syracuse" | "SYRACUSE"   
# "Rochester" | "ROCHESTER"
# "Albany" | "ALBANY"
#  "Ithaca" | "ITHACA" 

sort(unique(gra16$principal_place_state)) 
#Output is  "Illinois" "Massachusetts" "New York" "NEW YORK"
#this means that the filter in the manual download is not for this variable. 

######RECIPIENTS###############
sort(unique(gra16$recipient_state_code)) 
#output is only NY, so the filter is by recipients place, which makes sense because if we are an IDA we should care about the recipient not the pop.

head(sort(unique(gra16$recipient_city_name)))
# "Syracuse" | "SYRACUSE"   
# "Rochester" | "ROCHESTER"
# "Albany" | "ALBANY"
#  "Ithaca" | "ITHACA" 

head(sort(unique(gra16$recipient_county_code)))
head(sort(unique(gra16$recipient_county_name)))
head(sort(unique(gra16$recipient_state_code)))

```

#Issue with consistency of the names
```{r}
#looking for consistency in the names of agencies / programs

x <- gra16[, c("maj_agency_cat", "agency_name", "agency_code", "cfda_program_num", "cfda_program_title")]
x$maj <- substr(x$maj_agency_cat, 1,2)
x$cod <- substr(x$agency_code, 1,2)
x$cfda <- substr(as.character(x$cfda_program_num), 1,2)

#table(x$maj == x$cod) #only 4 inconsistencies between the first two numbers of agency maj and agency code.

#table(x$maj == x$cfda) #only 43 times they are the same
#and this is when:
#x[x$maj == x$cfda,] #19 for department of state in both cfda codes and agency code

length(unique(x$maj_agency_cat)) #24
length(unique(x$maj)) #23
length(unique(x$agency_code)) #83
length(unique(x$cod)) #25

length(unique(x$agency_name)) #67

length(unique(x$cfda)) #24
length(unique(x$cfda_program_num)) #782

```


#looking at action type

```{r}
#most common cfda

head(arrange(as.data.frame(table(gra16$cfda_program_title)), desc(Freq)))

x <- gra16$cfda_program_title == "Highway Planning and Construction"
dat <- gra16[x,]

#unique(gra16$action_type)
#A: new assistance
#B: continuation
#C: Revision
#D: Fnd adjustment

dat$act_typ <- substr(dat$action_type, 1, 1) #making it only the first code letter
head(arrange(as.data.frame(table(dat$act_typ)), desc(Freq)))

```



**ASSISSTANCE**

* 1	unique_transaction_id	character
* 16	federal_award_id	character

* 18	fed_funding_amount	numeric
* 19	non_fed_funding_amount	numeric
* 20	total_funding_amount	numeric

* 50	asst_cat_type	character
* 24	assistance_type	character
* 25	record_type	integer             :  Federal Assistance Awards Data System record type: 1 = county aggregate record, 2 = individual action record.
* 47	fiscal_year	integer

* 23	ending_date	character

* 28	principal_place_code	character
* 29	principal_place_state	character
* 30	principal_place_cc	character
* 32	principal_place_zip	character
* 48	principal_place_state_code	character

**AGENCY/PROGRAM**

* 52	maj_agency_cat	character
* 35	agency_name	character
* 4	  cfda_program_num	numeric
* 34	cfda_program_title	character
* 15	agency_code	character

**RECIPIENT**

* 36	project_description	character
* 7	recipient_name	character
* 13	recipient_type	character
* 49	recip_cat_type	character

* 8	recipient_city_code	integer
* 9	recipient_city_name	character
* 10	recipient_county_code	integer
* 11	recipient_county_name	character
* 56	recipient_state_code	character
* 12	recipient_zip	integer
* 42	receip_addr1	character
* 43	receip_addr2	character
* 44	receip_addr3	character

* 57	exec1_fullname	character
* 58	exec1_amount	numeric


#recip_cat_type and recipient_type equivalencies

* g: Government -> 00: State government           
* g: Government -> 02: City or township government
* g: Government -> 01: County government
* g: Government -> 04: Special district government
* g: Government -> 05: Independent school district
* h: Government -> 06: State controlled institution of higher education
 
* n: Nonprofit agencies -> "12: Other nonprofit"                                 
* n: Nonprofit agencies -> "11: Indian tribe"                                    

* h: Private agencies -> 20: Private higher education
* f: Private agencies -> 22: Profit organization
* f: Private agencies -> 23: Small business
* i: Private agencies -> 21: Individual
* o: Private agencies -> 25: All other


#Questions

1. how to work with negatives?

2. when looking at the action type 

3. Continuation (funding in succeeding budget period which stemmed from prior agreement to fund amount of the current action)"  This means we are double counting money?

 
                                


