---
title: "donuts yummy"
output: github_document
html_document:
  df_print: paged
  keep_md: true
  code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = T, message=F, warning=F, fig.width = 10, fig.height = 10 )
library(dplyr)
library(pander)
library(ggplot2)

########################
#grants DF

gra16 <- readRDS( gzcon(url("https://github.com/USAspendingexplorer/USAspending-explorer/blob/master/Data/Processed/NYgra16_cleaned.rds?raw=true")))

graTOTAL <- gra16

options(scipen = 999)
```


#USING PIE FUNCTION and exploring both total and no state grants for NY

##Main Federal Funding Agencies in NY

```{r}
#Possible variables:
#unique(gra16$maj_agency_cat) #27 levels
#unique(gra16$agency_name) #72 levels
#unique(gra16$agency_code) #95 levels


#grouping by Agencies (maj_agency_cat)
age <- group_by(gra16, maj_agency_cat)
age <- arrange(as.data.frame(summarize(age, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#grouping by Age TOTAL
ageT <- group_by(graTOTAL, maj_agency_cat)
ageT <- arrange(as.data.frame(summarize(ageT, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#barplot
par(mar = c(10,6,3,3), mfrow=c(1,2))
dat <- age
barplot(dat$Fed/1000000, names.arg = dat$maj_agency_cat, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8, col="#4979FF", main= "Main Federal Funding Agencies in NY")
dat <- ageT
barplot(dat$Fed/1000000, names.arg = dat$maj_agency_cat, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8, col="#4979FF", main = "Including State government")

#####pie charts!########

#making an others category
dat <- age
pie <- dat[1:5,]
pie[6,] <- c("OTHERS", sum(dat$Fed[6:length(dat$Fed)]), sum(dat$Freq[6:length(dat$Fed)])) 
pie$Fed <- as.numeric(pie$Fed)
pie <- arrange(pie, desc(Fed))

#making an others category TOTAL
dat <- ageT
pieT <- dat[1:5,]
pieT[6,] <- c("OTHERS", sum(dat$Fed[6:length(dat$Fed)]), sum(dat$Freq[6:length(dat$Fed)])) 
pieT$Fed <- as.numeric(pieT$Fed)
pieT <- arrange(pieT, desc(Fed))

#colors
col <- c("#4979FF", "#FF9858",  "#8DFFFF", "#FF84FF", "#55D473", "#FFFF78", "grey80")

#plotting pies
plot.new()
par(mfrow = c(1,2), mar=c(1,1,1,1))
pie <- pie
pie(as.numeric(pie$Fed), labels = paste0(pie$maj_agency_cat, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Main Federal Funding Agencies in NY", cex= .7, col=col)

pie<- pieT
pie(as.numeric(pie$Fed), labels = paste0(pie$maj_agency_cat, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Including State Government", cex= .7, col=col)

```


##Federal Funding by Recipient Types (recip_cat_type)

```{r}

#possible variables:
#unique(gra16$recip_cat_type) #12 levels
#unique(gra16$recipient_type) #7 levels

#grouping by recipients 
rec <- group_by(gra16, recip_cat_type)
rec <- arrange(as.data.frame(summarize(rec, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#grouping by recipients TOTAL
recT <- group_by(graTOTAL, recip_cat_type)
recT <- arrange(as.data.frame(summarize(recT, Fed = sum(fed_funding_amount), Freq = n())), desc(Fed))

#barplot
par(mar = c(7,6,3,3), mfrow=c(1,2))

barplot(rec$Fed/1000000, names.arg = rec$recip_cat_type, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8, col="#4979FF", main= "Federal Funding by Recipient Types")

barplot(recT$Fed/1000000, names.arg = recT$recip_cat_type, las = 2, cex.names = .7, cex.axis = .7 , ylab = "in USD 1MLL", cex.lab = .8, col="#4979FF", main = "Including State government")


#pie chart!
col <- c("#4979FF", "#FF9858",  "#8DFFFF", "#FF84FF", "#55D473", "#FFFF78", "grey80")

plot.new()
par(mfrow = c(1,2), mar=c(1,1,1,1))
pie <- rec
pie(as.numeric(pie$Fed), labels = paste0(pie$recip_cat_type, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Federal Funding by Recipient Types", cex= .7, col=col)

pie<- recT
pie(as.numeric(pie$Fed), labels = paste0(pie$recipient_name, "\n" , round(pie$Fed/sum(pie$Fed)*100, digits=1), "%"), main = "Including State Government", cex= .7, col=col)

```

#USING DONUTZZ FUNCTION: Major Agencies in NY State

First we need to create a function to read functions from github into Rstudio:
```{r}

#calling the function from git
source_github <- function( url ) {
  # load package
  require(RCurl)
  
  # read script lines from website and evaluate
  script <- getURL(url, ssl.verifypeer = FALSE)
  eval(parse(text = script), envir=.GlobalEnv)
} 

#load the donutzz function using the RAW link
source_github("https://raw.githubusercontent.com/icps86/Functions/master/krzydonutzz")
source_github("https://raw.githubusercontent.com/icps86/Functions/master/donutzz.R")
```

Now we need a data frame. 
```{r}
#We first need to make a DF
dat <- group_by(gra16, maj_agency_cat)
dat <- arrange(as.data.frame(summarize(dat, Fed = sum(fed_funding_amount))), desc(Fed))

#Removing code
dat$maj_agency_cat <- substr(dat$maj_agency_cat, 7,nchar(dat$maj_agency_cat))

#making an others category
dat[6,] <- c("Others", sum(dat$Fed[6:length(dat$Fed)])) 
dat <- dat[c(1:6),]
dat$Fed <- as.numeric(dat$Fed)
dat <- arrange(dat, desc(Fed))

#this is the dataframe we use
dat %>% pander

```

Now we can use the function to call the donut

```{r}
donutzz(x=dat$Fed, lev=dat$maj_agency_cat, main="Main Funding Agencies in NY State")
```

#KRZYDONUTZZ FUNCTION: Recipient types in 4 NY Counties

Making mutiple donuts with facet wrap function 

```{r}
#making dataset 
dat <- gra16
dat <- group_by(dat, Name, recip_cat_type)
dat <- arrange(as.data.frame(summarize(dat, Fed = sum(fed_funding_amount))), Name)
table(dat$Fed < 0) #has negative numbers
dat$Fed[0 >= dat$Fed] <- 0
table(is.na(dat$Fed)) #no NAs

x <- dat$Name%in% c("Albany", "Onondaga", "Wyoming", "Queens")
dat <- dat[x,]

#trying the function
krzydonutzz(x= dat, values = "Fed", labels = "recip_cat_type", multiple = "Name", main = "Federal Funding by Recipient type in NY Counties", percent.cex = 3, columns = 2, sbgfill = "grey95", sbgcol = "grey60")

```

#twiking the function
```{r}

x <- dat

krzydonutzz <- function (x, values, labels, multiple = NULL, width.max=3, width.min=2, border.col="white", border.cex =1.5,
                         main = "", mar = 1, percent.cex = 4, columns = 2, sbgfill = "grey95", sbgcol = "grey60")
{
  # Load Required Packages
  require(ggplot2)
  require(dplyr)
  
  #renaming the columns of the DF to work with them
  colnames(x)[colnames(x) == multiple] <- "multiple"
  colnames(x)[colnames(x) == labels] <- "labels"
  colnames(x)[colnames(x) == values] <- "values"
  
  #Warnings
  if (!is.numeric(x$values) || any(is.na(x$values) | x$values < 0))
    stop("'Values' variable cannot handle NA's and must be a numeric vector with only positive numbers.")
  
  if (length(unique(x$multiple)) > 8)
    stop("You are trying to plot too many donutzz! (8 max)")

  #making factors for labels
  x$labels <- factor(x=as.character(x$labels), ordered = F)

  #loop to format the dataframe and making the variables needed
  i <- 1
  df <- data.frame()
  x2 <- NULL
  
  while (i  <= length(unique(x$multiple))) 
  {
    x2 <- x$multiple == sort(unique(x$multiple))[i]
    x2 <- x[x2,]
    x2 <- x2[order(-x2$values),] 
    x2$fraction <- x2$values/sum(x2$values)
    x2$ymax <- cumsum(x2$fraction)
    x2$ymin <- c(0, x2$ymax[1:length(x2$ymax)-1])
    x2$mid <- (x2$ymin + x2$ymax)/2
    df <- rbind(df, x2)
    i = i + 1
  }
  
  x <- df
  
  pie <- ggplot(data=x,
                aes(fill=labels,
                    xmax=width.max, xmin=width.min,
                    ymax=ymax, ymin=ymin)) +
    xlim(c(0, width.max + mar)) +
    geom_rect(colour=border.col,
              size = border.cex) +
    coord_polar(theta="y") +
    geom_text(aes(x= width.max+.5, y = x$mid, label = paste0(round(x$fraction*100, digits = 1),"%")),
              size= percent.cex,
              #fontface = "bold",
              hjust = .5,
              vjust = .5,
              colour = "grey50") +
    ggtitle(label=main) +
    theme_bw() +
    theme(panel.grid=element_blank(),
          axis.text=element_blank(), 
          axis.ticks=element_blank(),
          axis.title = element_blank(),
          panel.border = element_blank(),
          strip.background = element_rect(fill= sbgfill, color = sbgcol), 
          plot.margin = margin(0,0,0,0, "cm"),
          legend.margin = margin(0,0,0,0, "cm"),
          legend.title = element_blank(),
          legend.position = "right",
          legend.text = element_text(colour = "grey40", size = 9)) +
    facet_wrap(~multiple, ncol = columns)
  return(pie)
}

krzydonutzz(x= dat, values = "Fed", labels = "recip_cat_type", multiple = "Name", main = "", percent.cex = 3, columns = 2, sbgfill = "grey95", sbgcol = "grey60")


```

